Active Learning from Blackbox to Timed Connectors

### 2016-04-03 shortpaper/paper.tex
.

\begin{document}
\maketitle 
\begin{abstract}
  Coordination models and languages play a key role in formally specifying the communication and
  interaction among different components in large-scale distributed and concurrent systems. In this
  paper, we propose an active learning framework to extract timed connector models from black-box
  system implementation. 
  We first introduce parameterized Mealy machine(PMM) as an operational semantic
  model for channel-based coordination language Reo. PMM serves as a bridge
  between Reo connectors and Mealy machines. With the product operator, complex connectors can be
  constructed by joining basic channels and the PMMs of connectors can be transformed into Mealy
  machines. Moreover, we adapt L*, a well-known learning algorithm, to timed connectors (in the form
  of Mealy machines). The new algorithm has shown its efficiency in multiple case studies. 
  This framework has been implemented in \texttt{Golang}.
\end{abstract}

\begin{IEEEkeywords}
  Active Learning, Coordination Languages, Timed Connectors
\end{IEEEkeywords}

@@

\emph{Timed Reo} is a real-time extension of the coordination language \emph{Reo}, which can be used
to describe the coordination process in DRES clearly and intuitively. Different formal
semantics have been proposed to specify the behavior of timed Reo.
For example, an operational semantics based on \emph{timed constraint automata} was raised by Baier et
al.\cite{DBLP:conf/sefm/ArbabBBR04}, where a variant of LTL was also proposed to describe the
properties of timed connectors.
In \cite{DBLP:conf/tase/Meng12}, a UTP design model was proposed to specify the behavior of timed
connectors..

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.4\textwidth]{./images/howto.png}
  \end{center}
  \caption{The Active Learning Framework}
  \label{fig:howto}
\end{figure}

@@

@@

\section{Preliminaries} 
\label{sec:preliminaries}
\subsection{Reo Coordination Language} 
\label{sec:reo}
We provide here a brief overview of the main concepts in Reo. Further details can be found in
\cite{DBLP:journals/mscs/Arbab04, DBLP:journals/scp/BaierSAR06}.

Reo is a channel based exogenous coordination language proposed by F. Arbab in
\cite{DBLP:journals/mscs/Arbab04}. 
A coordinator in Reo, also called \emph{connector}, provides the protocol that controls and
organizes the communication, synchronization and cooperation among the components which
communicate through the connector. Connectors can be defined without dependence on components,
which makes Reo a powerful ``glue language'' in component-based
development\cite{DBLP:journals/sigsoft/Gill03}.

In Reo, complex connectors are made up of simpler ones. The atomic connectors are called
\emph{channels}, and each of them has two \emph{channel ends}. There are two types of channel ends:
\emph{source} and \emph{sink}. Source channel ends accept data into the channel, while sink
channel ends dispense data out of the channel.

\begin{figure}[ht]
  \begin{center}
    \input{./images/basic.tex}
  \end{center}
  \caption{Basic Reo Channels}
  \label{fig:basic}
\end{figure}

Fig. \ref{fig:basic} shows the graphical representation of some basic channel types in Reo, whose
behavior are described as follows:

\begin{itemize}
  \item [-] A \emph{Sync} channel accepts a data item from its source end iff the data item can be
    dispensed through its sink end simultaneously.
  \item [-] A \emph{LossySync} channel is always ready to accept data items. These items will be
    dispensed through its sink end simultaneously if possible, otherwise they will be lost.
  \item [-] A \emph{SyncDrain} channel has two source ends and no sink end. It can accept a data
    item through one of its source end iff a data item is also available to be accepted
    simultaneously through the other end. Then both items will be dropped.
  \item [-] A \emph{FIFO1} channel is an asychronous channel with a buffer cell. It accepts a
    data item whenever the buffer is empty. 
  \item [-] A t-timer channel accepts any data item from its source end, and later dispense it to
    its sink end after a delay of $t$ time units.
    \footnote{
      Here the interpretation of t-Timer is a bit different as in other works like
      \cite{DBLP:conf/sefm/ArbabBBR04, DBLP:conf/tase/Meng12}, but we can easily transform between the
      two kinds of behavior by composing with some other basic channels.
      }
\end{itemize}

Channels are joint together in \emph{nodes}. There are three types of nodes:
\emph{source}, \emph{sink} and \emph{mixed} nodes, depending on whether all channel ends that
coincide on a node are source ends, sink ends or a combination of both.
(see Fig. \ref{fig:node})

\begin{figure}[ht]
  \begin{center}
    \input{./images/nodes.tex}
  \end{center}
  \caption{Source, Sink and Mixed Nodes in Reo}
  \label{fig:node}
\end{figure}

Components can be linked to source nodes or sink nodes. A component can write data items to its
corresponding source node only if the data item can be dispensed simultaneously to all source ends
on this node. Similarly, a component can read a data item only if there is at least one readable
sink end on its corresponding node. A mixed node non-determinstically selects and takes a data item
from one of its coincident sink ends and replicates it into all of its coincident source ends.

\begin{figure}[ht]
  \begin{center}
    \input{./images/connectorexample.tex}
  \end{center}
  \caption{An Example of Coordination with Reo Connectors}
  \label{fig:reoconnector}
\end{figure}

Fig. \ref{fig:reoconnector} provides a simple example to illustrate how complex
connectors are constructed and used in coordination. In this example, we have three components
COM1,COM2 and MONITOR. With an \emph{alternator} connector as shown in Fig.
\ref{fig:reoconnector}, MONITOR can obtain data items from COM1 and COM2 alternately. This example
has been implemented in \texttt{Golang} as shown in Section \ref{sec:reolib}.
Besides, all the connectors are supposed to have \emph{deterministic} behavior hereinafter to meet
the requirement of active learning algorithm L*.

\subsection{Mealy Machines}

In this work, we use \emph{Mealy machines} to model Reo connectors.
Mealy machine was first proposed by George. H. Mealy in \cite{George1955A} as an extension of finite
state machine. Compared with other variants, Mealy machines are designed to model
reactive systems, where outputs of a machine are determined by not only its current state but also
the current inputs. Besides, Mealy machines are supposed to be \emph{input enabled}, which means
that all possible inputs should be acceptable in all states. In other words, if an input is invalid
for some state, we need to manually define an additional state to describe such exceptions.

Various forms of Mealy machines have been defined in the
literature\cite{George1955A,DBLP:journals/jcss/Broy14,DBLP:conf/sfm/SteffenHM11}. 
Since active automata learning requires the system-under-learn to be deterministic, in
this paper we take the deterministic Mealy machine model as defined in 
\cite{DBLP:conf/sfm/SteffenHM11}.

\begin{definition}[Mealy machine]
  A Mealy machine is a 6-tuple $(S, s_0, I, O, \delta, \lambda)$ consisting of:
  \begin{itemize}
    \item[-] a finite set of states $S$,
    \item[-] an initial state $s_0\in S$,
    \item[-] a finite set of inputs $I$,
    \item[-] a finite set of outputs $O$,
    \item[-] an output function $\delta :  S \times  I \rightarrow O$ mapping a pair
      of a state and an input to the corresponding output,
    \item[-] a transition function $\lambda : S \times I \rightarrow S$ mapping a pair of a
      state and an input to the corresponding successor state.
  \end{itemize}
\end{definition}

We say that a Mealy machine is \emph{finite} if the set $S$ of states and the set $I$ of inputs are
both finite. Hereinafter, Mealy machines are always supposed to be finite in this paper.

\subsection{Active Learning}
In this section, we briefly introduce the main ideas of active automata learning.

Active learning \cite{settles2010active} is a special case of semi-supervised machine learning where
a learning algorithm is able to interactively query the target systems to obtain the desired outputs
under certain inputs. By well-designed query strategy, active learning is able to obtain more
accurate models with smaller dataset.

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.5\textwidth]{./images/activelearning.png}
  \end{center}
  \caption{Active Automata Learning}
  \label{fig:activelearning}
\end{figure}

Fig. \ref{fig:activelearning} shows the sketch of active learning, wherein:
\begin{itemize}
  \item[-] \emph{Teacher} and \emph{Student}: Active learning is an interactive process where
    students ask questions and teachers answer. Here learning algorithm plays the role of student.
  \item[-] \emph{Oracles} is an interface specifying which kind of questions can be answered by the
    teacher.
  \item[-] \emph{SUL} is an abbreviation of System Under Learn. In this paper, we take blackbox
    models as our SULs.
  \item[-] \emph{CEGAR} indicates Counter-Example Guided Abstraction
    Refinement\cite{DBLP:conf/cav/ClarkeGJLV00}. In active learning, we need counter-examples to
    guide us on further queries and cover the undistinguished states.
\end{itemize}

When applying active automata learning on some model, we assume that the model should be equivalent
to some \emph{Mealy machine}, i.e., a deterministic model accepting a finite set of inputs and
mapping them to a finite set of outputs.

Such a model is encapsulated as a \emph{teacher} by the \emph{Oracle}, which handles
all communication with the model. The oracle also serves as a so-called \emph{Minimal Adequate Teacher}
interface, which is responsible for two types of queries.

\begin{itemize}
  \item[-] \textbf{Membership Query} (hereinafter referred to as \emph{mq}) In grammar-learning
    \cite{DBLP:journals/iandc/Angluin87}, \emph{mq} checks if a word is a member of certain language
    defined by the given grammar. When it comes to
    automata learning, \emph{mq} is supposed to provide \emph{simulation results} for given input
    sequences.
  \item[-] \textbf{Equivalence Query} (hereinafter referred to as \emph{eq}) Given a hypothesis
    (usually constructed by the learning algorithm), \emph{eq} checks whether the hypothesis is
    equivalent to the system-under-learn and generates a counter-example if needed. Generally, 
    \emph{eq} is irrealizable when SUL is a blackbox. So we tactically use $mq$ to achieve the
    approximate results.
\end{itemize}

These queries are given by \emph{learning algorithms}, or so-called \emph{students} in Fig.
\ref{fig:activelearning}. From the \emph{mq} results, a learning algorithm constructs a
\emph{hypothesis} and then check it with \emph{eq}. If counter-examples are found, we turn back and
repeat the hypothesis construction until the equivalence query returns \emph{true}.

More details on the active automata learning algorithm will be presented in Section
\ref{sec:activelearning}.

@@

\subsection{Time Domain}
Time has been investigated in several extensions of Reo. For example, timed
Reo\cite{DBLP:conf/sefm/ArbabBBR04, DBLP:conf/fmoods/MengA07}, hybrid Reo\cite{DBLP:conf/icfem/ChenSS14}, etc.
Generally, these models are designed to handle real-time behavior where time is defined on the
real number field $\mathbb{R}$. However, there are also some works like
\cite{DBLP:journals/fmsd/PrabhakarDM015} where
rational time indeed  makes things easier. In this paper, we choose the rational number field
$\mathbb{Q}$ as our time domain, which simplifies discretization of timed behavior greatly.

As presented in section \ref{sec:reo}, all real-time behavior in timed Reo comes with a finite set
of timed channels. We use
$t_i\in\mathbb{Q}, i=1,2,\dots,n$ to denote the delays of these timed channels, and then we can define a precision
function $prec$ which calculate a gcd-style \emph{maximal time precision}:
\[
prec(t_1,\dots,t_n) = \max\{T\in\mathbb{Q}|\forall t_i.\exists n_i\in\mathbb{N}.t_i=n_i\cdot T\}
\]
It's easy to prove that such a $T$ always exists.

In embedded systems, the maximal time precision is also called ``clock-period'', which is the basic
time unit provided by an oscillator. For example, a widely-used Intel-80C51 microcontroller works
under the frequency 16Mhz, where the clock-period is $62.5$ns. A reasonable assumption comes that
such a maximal time precision $dt=prec(t_1,\dots,t_n)$ should always be provided to make sure
different components are able to work together. Then all $t_i$-Timers can be seen as $n_idt$-Timers
for some $n_i$ and we use $n_i$-Timers instead in the following sections.

Besides, we add a timed action ``$T$'' in Mealy machines. It indicates that the corresponding
transition will take one time unit to finish.

@@

\begin{example}[FIFO1]
  \label{example:pmmfifo}
  The PMM of a FIFO1 with source end $A$ and sink end $B$ can be defined as follows.
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_0\}\cup\{q_d|d\in\Sigma\}$
    \item[-] $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    \item[-] output function
      \begin{small}
        \begin{displaymath}
          \delta(\Sigma)(s,i)=\left\{
          \begin{array}[ht]{ll}
            \rempty & s=q_0\land i=\{A\mapsto\_,B\mapsto\rnoread\} \\
            \rblock & s=q_0\land i=\{A\mapsto d',B\mapsto\rread\} \\
            \rblock & s=q_0\land i=\{A\mapsto\bot,B\mapsto\rread\} \\
            \rempty & s=q_d\land i=\{A\mapsto\bot,B\mapsto\rnoread\} \\     
            \{B\mapsto d\} & s=q_d\land i=\{A\mapsto\_,B\mapsto\rread\}\\
            \rblock & s=q_d\land i=\{A\mapsto d',B\mapsto \rnoread\} \\
          \end{array}
          \right.
        \end{displaymath}
      \end{small}
    \item[-] transition function
      \begin{small}
        \begin{displaymath}
          \lambda(\Sigma)(s,i)=\left\{
          \begin{array}[ht]{ll}
            q_{d'} & s=q_0\land i=\{A\mapsto d',B\mapsto\_\} \\
            q_{d'} & s=q_{d}\land i=\{A\mapsto d',B\mapsto \rread\} \\
            q_{d} & s=q_{d}\land i=\{A\mapsto \_,B\mapsto \rnoread\} \\
            q_0 & otherwise \\
          \end{array}
          \right.
        \end{displaymath}
      \end{small}
  \end{itemize}
\end{example}

Besides, the concrete Mealy machine (where $\Sigma=\{d\}$) is shown in Fig. \ref{fig:pmmfifo},
where labels of edges are in form of $\frac{input}{output}$. All $\rblock$ edges are ignored for
clearance.
\begin{figure}[ht]
  \begin{center}
    \input{./images/pmmfifo.tex}
  \end{center}
  \caption{PMM-based Semantics of $\smap{FIFO1}_\Sigma$, where $\Sigma=\{d\}$}
  \label{fig:pmmfifo}
\end{figure}

\begin{example}[Sync]
  The PMM of a Sync with a source-end A and a sink-end B can be defined as:
  \begin{itemize}
    \item[-] $S(\Sigma)=\{q_0\}$, $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    \item[-] output function
      \begin{displaymath}
        \delta(\Sigma)(s,i)=\left\{
        \begin{array}[ht]{ll}
          \rempty & i=\{A\mapsto\bot,B\mapsto\rnoread\} \\
          \{B\mapsto d\} & i=\{A\mapsto d,B\mapsto\rread\}\\
          \rblock & i=\{A\mapsto\bot,B\mapsto\rread\} \\
          \rblock & i=\{A\mapsto d,B\mapsto\rnoread\} \\
        \end{array}
        \right.
      \end{displaymath}
    \item[-] transition function $\lambda(\Sigma)(s,i)=q_0$.
  \end{itemize}
\end{example}

@@

\begin{figure}[ht]
  \begin{center}
    \begin{displaymath}
      \begin{array}{l||ccccc||c}
        \hline
        & (a,\rread) & (\bot,\rread) & (a,\rnoread) & (\bot,\rnoread) & T & acc\\
        \hline\hline
        \varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        \hline
        (a,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        (\bot,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        \rowcolor[gray]{.9}
        (a,\rnoread) & \rblock & \rblock & \rblock & \rempty & \rempty & unclosed\\
        (\bot,\rnoread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        T & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        \hline
      \end{array}
    \end{displaymath}
    % \input{./images/hypo.tex}
  \end{center}
  \caption{Observation Table and Corresponding Hypothesis}
  \label{fig:hypo}
\end{figure}

@@

\begin{figure}[ht]
  \begin{center}
    \begin{displaymath}
      \begin{array}{l||ccccc||c}
        \hline
        & (a,\rread) & (\bot,\rread) & (a,\rnoread) & (\bot,\rnoread) & T & acc\\
        \hline\hline
        \varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        (a,\rnoread) & \rblock & \rblock & \rblock & \rempty & \rempty & (a,\rnoread) \\
        \hline
        (a,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        (\bot,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        (\bot,\rnoread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        T & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        (a,\rnoread),(a,\rread) & \rblock & \rblock & \rblock & \rempty & \rempty & (a,\rnoread) \\
        (a,\rnoread),(\bot,\rread) & \rblock & \rblock & \rblock & \rempty & \rempty & (a,\rnoread) \\
        (a,\rnoread),(a,\rnoread) & \rblock & \rblock & \rblock & \rempty & \rempty & (a,\rnoread) \\
        (a,\rnoread),(\bot,\rnoread) & \rblock & \rblock & \rblock & \rempty & \rempty &
        (a,\rnoread) \\
        (a,\rnoread),T & \rblock & \rblock & \rblock & \rempty & \rempty & (a,\rnoread) \\
        \hline
      \end{array}
    \end{displaymath}
  \end{center}
  \caption{A Closed Version of Fig. \ref{fig:hypo}}
  \label{fig:hypo2}
\end{figure}

@@

\subsection{Performance Optimization}
As a well-known learning algorithm, L* has proved its efficiency in models without time.
However, when dealing with timed connectors, the algorithm failed to meet our expectation.

\begin{table}[ht]
  \renewcommand{\arraystretch}{1.3}
  \caption{Time-Cost Analysis}
  \label{tabel:timecost}
  \centering
  \begin{tabular}{l||rrr}
    \hline
    & FIFO1 & Alternator & Gate \\
    \hline\hline
    Membership Query(s) & 41.571 & 126.468 & 169.161 \\
    Hypothesis Query(s) & 0.001 & 0.003 & 0.004 \\
    Total Time(s) & 41.715 & 165.114 & 247.098 \\
    Membership Query(\%) & 99.6 & 76.6 & 68.5 \\
    \hline
  \end{tabular}
\end{table}

@@

\begin{table}[ht]
  \renewcommand{\arraystretch}{1.3}
  \caption{Reduction of Membership Queries}
  \label{tabel:cacheoptimization}
  \centering
  \begin{tabular}{l||rrr}
    \hline
    & FIFO & Alternator & Gate \\
    \hline\hline
    Original Algorithm & 93 & 880 & 1034 \\
    Cached Algorithm & 90 & 725 & 707 \\
    Reduction Rate & 3.2\% & 21.4\% & 31.6\% \\
    \hline
  \end{tabular}
\end{table}

@@

\subsection{An Example of the Reo Package }
\label{sec:reolib}

Our implementation in \texttt{Golang} is well-prepared not only for academic use
but also for practical concurrent programming. The following code shows how to construct an alternator
connector (see Fig. \ref{fig:reoconnector}) in \texttt{Golang}. An intact version of this example
can be found in our github repo.

\begin{lstlisting}
func alternator(A, B, C Port) {
  M := MakePorts(6)

  // definition of channels
  go ReplicatorChannel(A, M[0], M[1])
  go ReplicatorChannel(B, M[2], M[3])
  go MergerChannel(M[4], M[5], C)

  go SyncdrainChannel(M[1], M[2])
  go SyncChannel(M[0], M[4])
  go FifoChannel(M[3], M[5])
}
\end{lstlisting}

Provided with the source and sink nodes, \emph{alternator} function creates a series of basic
channels and mixed nodes (named \emph{Port}) to serve as the alternator connector we need. Now we
can activate the components and using \emph{alternator} function to coordinate them.

\begin{lstlisting}
A,B,C := MakePorts(3)
alternator(A, B, C)

go sender(A, "MSG A")
go sender(B, "MSG B")
go monitor(C)
\end{lstlisting}

In this case, \emph{senders} are goroutines (basic parallel units in \texttt{Golang}) that keep
sending certain messages to some given port (A and B). A \emph{monitor} keeps trying to read data
items from the sink end C and print them on the screen. Finally, we have an interleaved sequence of
``MSG\_A'' and ``MSG\_B''.

@@

We also adapt the well-known active learning algorithm L* to deal with time domain.
When time is taken into consideration, the original L* algorithm run into a performance bottleneck.
By a tree-style cache, we make significant reduction on membership queries and, in turn,
improve the performance of learning algorithm. As a by-product, we also encapsulate the Reo
connector as a distributable package which could contribute to concurrent programming.

>>>>

\begin{document}
\maketitle 
\begin{abstract}
  Coordination models and languages play a key role in formally specifying the communication and
  interaction among different components in large-scale distributed and concurrent systems. In this
  paper, we use active learning to extract timed connector models from black-box
  system implementation. 
  We first introduce parameterized Mealy machine(PMM) as an operational semantic
  model for channel-based coordination language Reo. PMM serves as a bridge
  between Reo connectors and Mealy machines. With the product operator, complex connectors can be
  constructed by joining basic channels and the PMMs of connectors can be transformed into Mealy
  machines. Moreover, we adapt L*, a well-known learning algorithm, to timed connectors (in the form
  of Mealy machines). The new algorithm has shown its efficiency in multiple case studies. 
\end{abstract}

\begin{IEEEkeywords}
  Active Learning, Coordination, Connectors
\end{IEEEkeywords}

@@

\emph{Timed Reo} is a real-time extension of the coordination language \emph{Reo}, which can be used
to describe the coordination process in DRES clearly and intuitively. 
\todo{describe semantics in a more short way}
%Different formal semantics have been proposed to specify the behavior of timed Reo.
%For example, an operational semantics based on \emph{timed constraint automata} was raised by Baier et
%al.\cite{DBLP:conf/sefm/ArbabBBR04}, where a variant of LTL was also proposed to describe the
%properties of timed connectors.
%In \cite{DBLP:conf/tase/Meng12}, a UTP design model was proposed to specify the behavior of timed
%connectors..

@@

%\begin{figure}[ht]
  %\begin{center}
    %\includegraphics[width=.4\textwidth]{./images/howto.png}
  %\end{center}
  %\caption{The Active Learning Framework}
  %\label{fig:howto}
%\end{figure}

@@

\todo{use a short paragraph to describe the preliminaries}

@@

@@

\todo{time domain?}

%\subsection{Time Domain}
%Time has been investigated in several extensions of Reo. For example, timed
%Reo\cite{DBLP:conf/sefm/ArbabBBR04, DBLP:conf/fmoods/MengA07}, hybrid Reo\cite{DBLP:conf/icfem/ChenSS14}, etc.
%Generally, these models are designed to handle real-time behavior where time is defined on the
%real number field $\mathbb{R}$. However, there are also some works like
%\cite{DBLP:journals/fmsd/PrabhakarDM015} where
%rational time indeed  makes things easier. In this paper, we choose the rational number field
%$\mathbb{Q}$ as our time domain, which simplifies discretization of timed behavior greatly.

%As presented in section \ref{sec:reo}, all real-time behavior in timed Reo comes with a finite set
%of timed channels. We use
%$t_i\in\mathbb{Q}, i=1,2,\dots,n$ to denote the delays of these timed channels, and then we can define a precision
%function $prec$ which calculate a gcd-style \emph{maximal time precision}:
%\[
%prec(t_1,\dots,t_n) = \max\{T\in\mathbb{Q}|\forall t_i.\exists n_i\in\mathbb{N}.t_i=n_i\cdot T\}
%\]
%It's easy to prove that such a $T$ always exists.

%In embedded systems, the maximal time precision is also called ``clock-period'', which is the basic
%time unit provided by an oscillator. For example, a widely-used Intel-80C51 microcontroller works
%under the frequency 16Mhz, where the clock-period is $62.5$ns. A reasonable assumption comes that
%such a maximal time precision $dt=prec(t_1,\dots,t_n)$ should always be provided to make sure
%different components are able to work together. Then all $t_i$-Timers can be seen as $n_idt$-Timers
%for some $n_i$ and we use $n_i$-Timers instead in the following sections.

%Besides, we add a timed action ``$T$'' in Mealy machines. It indicates that the corresponding
%transition will take one time unit to finish.

@@

%\begin{example}[FIFO1]
  %\label{example:pmmfifo}
  %The PMM of a FIFO1 with source end $A$ and sink end $B$ can be defined as follows.
  %\begin{itemize}
    %\item[-] $S(\Sigma)=\{q_0\}\cup\{q_d|d\in\Sigma\}$
    %\item[-] $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    %\item[-] output function
      %\begin{small}
        %\begin{displaymath}
          %\delta(\Sigma)(s,i)=\left\{
          %\begin{array}[ht]{ll}
            %\rempty & s=q_0\land i=\{A\mapsto\_,B\mapsto\rnoread\} \\
            %\rblock & s=q_0\land i=\{A\mapsto d',B\mapsto\rread\} \\
            %\rblock & s=q_0\land i=\{A\mapsto\bot,B\mapsto\rread\} \\
            %\rempty & s=q_d\land i=\{A\mapsto\bot,B\mapsto\rnoread\} \\     
            %\{B\mapsto d\} & s=q_d\land i=\{A\mapsto\_,B\mapsto\rread\}\\
            %\rblock & s=q_d\land i=\{A\mapsto d',B\mapsto \rnoread\} \\
          %\end{array}
          %\right.
        %\end{displaymath}
      %\end{small}
    %\item[-] transition function
      %\begin{small}
        %\begin{displaymath}
          %\lambda(\Sigma)(s,i)=\left\{
          %\begin{array}[ht]{ll}
            %q_{d'} & s=q_0\land i=\{A\mapsto d',B\mapsto\_\} \\
            %q_{d'} & s=q_{d}\land i=\{A\mapsto d',B\mapsto \rread\} \\
            %q_{d} & s=q_{d}\land i=\{A\mapsto \_,B\mapsto \rnoread\} \\
            %q_0 & otherwise \\
          %\end{array}
          %\right.
        %\end{displaymath}
      %\end{small}
  %\end{itemize}
%\end{example}

%Besides, the concrete Mealy machine (where $\Sigma=\{d\}$) is shown in Fig. \ref{fig:pmmfifo},
%where labels of edges are in form of $\frac{input}{output}$. All $\rblock$ edges are ignored for
%clearance.
%\begin{figure}[ht]
  %\begin{center}
    %\input{./images/pmmfifo.tex}
  %\end{center}
  %\caption{PMM-based Semantics of $\smap{FIFO1}_\Sigma$, where $\Sigma=\{d\}$}
  %\label{fig:pmmfifo}
%\end{figure}

%\begin{example}[Sync]
  %The PMM of a Sync with a source-end A and a sink-end B can be defined as:
  %\begin{itemize}
    %\item[-] $S(\Sigma)=\{q_0\}$, $I=\{A\}$, $O=\{B\}$, $s_0=q_0$
    %\item[-] output function
      %\begin{displaymath}
        %\delta(\Sigma)(s,i)=\left\{
        %\begin{array}[ht]{ll}
          %\rempty & i=\{A\mapsto\bot,B\mapsto\rnoread\} \\
          %\{B\mapsto d\} & i=\{A\mapsto d,B\mapsto\rread\}\\
          %\rblock & i=\{A\mapsto\bot,B\mapsto\rread\} \\
          %\rblock & i=\{A\mapsto d,B\mapsto\rnoread\} \\
        %\end{array}
        %\right.
      %\end{displaymath}
    %\item[-] transition function $\lambda(\Sigma)(s,i)=q_0$.
  %\end{itemize}
%\end{example}

@@

%\begin{figure}[ht]
  %\begin{center}
    %\begin{displaymath}
      %\begin{array}{l||ccccc||c}
        %\hline
        %& (a,\rread) & (\bot,\rread) & (a,\rnoread) & (\bot,\rnoread) & T & acc\\
        %\hline\hline
        %\varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        %\hline
        %(a,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        %(\bot,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        %\rowcolor[gray]{.9}
        %(a,\rnoread) & \rblock & \rblock & \rblock & \rempty & \rempty & unclosed\\
        %(\bot,\rnoread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        %T & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        %\hline
      %\end{array}
    %\end{displaymath}
    %% \input{./images/hypo.tex}
  %\end{center}
  %\caption{Observation Table and Corresponding Hypothesis}
  %\label{fig:hypo}
%\end{figure}

@@

%\begin{figure}[ht]
  %\begin{center}
    %\begin{displaymath}
      %\begin{array}{l||ccccc||c}
        %\hline
        %& (a,\rread) & (\bot,\rread) & (a,\rnoread) & (\bot,\rnoread) & T & acc\\
        %\hline\hline
        %\varepsilon & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        %(a,\rnoread) & \rblock & \rblock & \rblock & \rempty & \rempty & (a,\rnoread) \\
        %\hline
        %(a,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon\\
        %(\bot,\rread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        %(\bot,\rnoread) & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        %T & \rblock & \rblock & \rempty & \rempty & \rempty & \varepsilon \\
        %(a,\rnoread),(a,\rread) & \rblock & \rblock & \rblock & \rempty & \rempty & (a,\rnoread) \\
        %(a,\rnoread),(\bot,\rread) & \rblock & \rblock & \rblock & \rempty & \rempty & (a,\rnoread) \\
        %(a,\rnoread),(a,\rnoread) & \rblock & \rblock & \rblock & \rempty & \rempty & (a,\rnoread) \\
        %(a,\rnoread),(\bot,\rnoread) & \rblock & \rblock & \rblock & \rempty & \rempty &
        %(a,\rnoread) \\
        %(a,\rnoread),T & \rblock & \rblock & \rblock & \rempty & \rempty & (a,\rnoread) \\
        %\hline
      %\end{array}
    %\end{displaymath}
  %\end{center}
  %\caption{A Closed Version of Fig. \ref{fig:hypo}}
  %\label{fig:hypo2}
%\end{figure}

@@

\subsection{Performance Optimization}
As a well-known learning algorithm, L* has proved its efficiency in models without time.
However, when dealing with timed connectors, the algorithm failed to meet our expectation.
\todo{draw the two tables in github}
%\begin{table}[ht]
  %\renewcommand{\arraystretch}{1.3}
  %\caption{Time-Cost Analysis}
  %\label{tabel:timecost}
  %\centering
  %\begin{tabular}{l||rrr}
    %\hline
    %& FIFO1 & Alternator & Gate \\
    %\hline\hline
    %Membership Query(s) & 41.571 & 126.468 & 169.161 \\
    %Hypothesis Query(s) & 0.001 & 0.003 & 0.004 \\
    %Total Time(s) & 41.715 & 165.114 & 247.098 \\
    %Membership Query(\%) & 99.6 & 76.6 & 68.5 \\
    %\hline
  %\end{tabular}
%\end{table}

@@

%\begin{table}[ht]
  %\renewcommand{\arraystretch}{1.3}
  %\caption{Reduction of Membership Queries}
  %\label{tabel:cacheoptimization}
  %\centering
  %\begin{tabular}{l||rrr}
    %\hline
    %& FIFO & Alternator & Gate \\
    %\hline\hline
    %Original Algorithm & 93 & 880 & 1034 \\
    %Cached Algorithm & 90 & 725 & 707 \\
    %Reduction Rate & 3.2\% & 21.4\% & 31.6\% \\
    %\hline
  %\end{tabular}
%\end{table}

@@

@@

We also adapt the well-known active learning algorithm L* to deal with time domain.
% When time is taken into consideration, the original L* algorithm run into a performance bottleneck.
By a tree-style cache, we make significant reduction on membership queries and, in turn,
improve the performance of learning algorithm. As a by-product, we also encapsulate the Reo
connector as a distributable package which could contribute to concurrent programming.
